---
title: "WChange"
author: "NellySelem"
date: "December 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Whats change genomic dynamics



```{r cars}
library(ggplot2)
library( reshape )
library(plyr )
library(RColorBrewer)
families <- read.table("100Archea/100Archaea.r.allwc", header=TRUE, sep="\t")
hm.palette <- colorRampPalette(rev(brewer.pal(11, 'Spectral')), space='Lab')  
Frecuencia=sort(table(families$FunctionID),decreasing = TRUE)
head(Frecuencia)
Frecuencia[1]
indx <- which((Frecuencia[]) > 110) ;
#Name=families[which(families$FunctionID==indx),]
#table(Name$Function)
indx
```
##Data Bases
-Actino  

-Archaea 

-Cyanos  

In order to investigate some other functions  we took 100Cyanos, 100Archaea , 118 Actinos Closed  
Those families present on at least every organism  to asure they have expanssions  
Cyanos  
Actinos  
Archaea  
Streptomyces  
Those families expanded on at least two groups  

Those Families expanded on Archaea  and not expanded on Actino  

Those Families expanded on Actino and not on Archaea  
Those Families expanded on Cyano   

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
