---
header-includes:
- \usepackage{graphicx,latexsym}
- \usepackage{amssymb,amsthm,amsmath}
- \usepackage{longtable,booktabs,setspace}
---

<!--- Cyanobacteria-->

```{r include_packages_5, include = FALSE}
# This chunk ensures that the reedtemplates package is
# installed and loaded. This reedtemplates package includes
# the template files for the thesis and also two functions
# used for labeling and referencing
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(dplyr))
    install.packages("dplyr", repos = "http://cran.rstudio.com")
if(!require(ggplot2))
    install.packages("ggplot2", repos = "http://cran.rstudio.com")
if(!require(reedtemplates)){
  library(devtools)
  devtools::install_github("ismayc/reedtemplates")
  }
library(reedtemplates)
flights <- read.csv("data/flights.csv")
```


# Cyanobacteria EvoMining Results {#ref_labels}
Cyanobacteria is phylum {Referencia de luis}  

## Tables

----------------------------------------------------------------------------------
  Factors                    Correlation between Parents & Child     
------------------------- ----------------------------------------- --------------
  GenomeDB                                  1245                         
  
  Families                                  65                          
------------------------- ----------------------------------------- --------------
Table: Families on Cyanobacteria \label{tab:inher}

\clearpage
<!-- clearpage ends the page, and also dumps out all floats.Floats are things like tables and figures. -->

```{r loading_tables_5}
library(ggplot2)
library( reshape )
library(plyr )
library(RColorBrewer)
#setwd("/home/nelly/GIT/phd/Thesis/tesis/chap5_files")
CyanosCentral <- read.table("chapter5/CyanosCentral", header=TRUE, sep="\t")
CyanosHeatPlot <- read.table("chapter5/CyanosHeatPlot", header=TRUE, sep="\t")
CyanosNp <- read.table("chapter5/CyanosNp", header=TRUE, sep="\t")
CyanosSMASH <- read.table("chapter5/CyanosSMASH", header=TRUE, sep="\t")
CyanosTaxa <- read.table("chapter5/CyanosTaxa", header=TRUE, sep="\t")
```


## Cyanobacterias AntiSMASH 
```{r}
#CyanosHeatPlot$RastId <- with(CyanosHeatPlot, reorder(CyanosHeatPlot$RastIds, CyanosHeatPlot$TOTAL))
#ggplot(CyanosHeatPlot, aes(variable, RastId))+ geom_raster(aes(fill=Value)) + theme(text = element_text(size=8), axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +coord_equal()+scale_fill_gradient(low = "white", high = "black",na.value = "orange")
```

## Figures
Lets analize Phosphoribosyl_isomerase_3 family  
 
If your thesis has a lot of figures, _R Markdown_ might behave better for you than that other word processor.  One perk is that it will automatically number the figures accordingly in each chapter.    You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier.  If you label your figures, you can move the figures around and _R Markdown_ will automatically adjust the numbering for you.  No need for you to remember!  So that you don't have to get too far into \LaTeX\ to do this, a couple **R** functions have been created for you to assist.  You'll see their use below.

<!--
One thing that may be annoying is the way _R Markdown_ handles "floats" like tables and figures (it's really \LaTeX's fault). \LaTeX\ will try to find the best place to put your object based on the text around it and until you're really, truly done writing you should just leave it where it lies. There are some optional arguments specified in the options parameter of the `label` function.  If you need to shift your figure around, it might be good to look here on tweaking the options argument:  <https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions>

If you need a graphic or tabular material to be part of the text, you can just put it inline. If you need it to appear in the list of figures or tables, it should be placed in a code chunk.
-->


In the **R** chunk below, we will load in a picture stored as `reed.jpg` in our main directory.  We then give it the caption of "Reed logo", the label of "reed", and specify that this is a figure.  Note again the use of the `results = "asis"` specification to automatically include and compile the \LaTeX\ code.

```{r reed_logo3, results = "asis"}
label(path = "figure/reed.jpg", caption = "Reed logo", 
      label = "reed", type = "figure")
```

Here is a reference to the Reed logo: `r ref("reed")`.  Note the use of the inline **R** code here.  By default "figure" is specified as the type.  For clarity, we could have also added the `label` and `type` to the parameter specifications and this would give us the same result: `r ref(label = "reed", type = "figure")`.

\clearpage 

<!-- starts a new page and stops trying to place floats such as tables and figures -->

Below we will investigate how to save the output of an **R** plot and label it in a way similar to that done above.  Recall the `flights` dataset from [](#rmd-basics).  (Note that we've shown a different way to reference a section or chapter here.)  We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.  Note also the use of the `scale` parameter which is discussed on the next page.

```{r boxplot_3, warnings = FALSE, messages = FALSE}
delay_airline <- flights %>% group_by(carrier) %>%
  summarize(mean_dep_delay = mean(dep_delay)) %>%
  ggplot(aes(x = carrier, y = mean_dep_delay)) +
  geom_bar(position = "identity", stat = "identity", fill = "red")
ggsave("figure/delays.pdf", plot = delay_airline,
  height = 3, width = 6)
```

```{r box_label_3, results = "asis"}
label(path = "figure/delays.pdf", 
      caption = "Mean Delays by Airline", 
      label = "delays", type = "figure")
```

A table linking these carrier codes to airline names is available at <https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv>.

\clearpage

Next, we will explore the use of the `scale` parameter which can be used to shrink or expand an image.  Here we use the mathematical graph stored in the "subdivision.pdf" file.  Note that we didn't specify the `caption =` or `label =` here, but we could have.

```{r subdiv1_3, results = "asis"}
label("figure/subdivision.pdf", "Subdiv. graph", "subd", 
      scale = 0.75)
```

Here is a reference to this image: `r ref("subd")`.  (Move this around throughout the document as you wish.)

#### More Figure Stuff

Lastly, we will explore how to rotate figures using the `angle` parameter.

```{r subdiv1_180_3, results = "asis"}
label("figure/subdivision.pdf", 
      "A Larger Figure, Flipped Upside Down", 
      scale = 1.5,
      angle = 180,
      label = "subd2")
```

As another example, here is a reference to this figure: `r ref("subd2", type = "figure")`.  

#### Common Modifications

The following figure features the more popular changes thesis students want to make to their figures. We can add math to the caption that displays below the picture, specify the size of our caption to display below the figure (list of sizes available at this [link](http://www.emerson.emory.edu/services/latex/latex_169.html#SEC169)), and also specify that a different caption `alt.cap` be what appears in the Table of Figures for this figure.

If you'd like to make further tweaks to figures, you might need to invoke some \LaTeX\ code.  Please email us at <data@reed.edu> if you need assistance.

```{r tinymathcap_4, results = "asis"}
label("figure/subdivision.pdf", 
      caption = "Subdivision of arc segments",
      alt.cap = "You can see that $p_3 = p_6^\\prime$",
      cap.size = "footnotesize",
      label = "subd3")
```

<!-- Note here that you need to have two backslashes before LaTeX commands such as prime -->

## Footnotes and Endnotes

You might want to footnote something. ^[footnote text] The footnote will be in a smaller font and placed appropriately. Endnotes work in much the same way. More information can be found about both on the CUS site or feel free to reach out to <data@reed.edu>.

## Bibliographies

Of course you will need to cite things, and you will probably accumulate an armful of sources. There are a variety of tools available for creating a bibliography database (stored with the .bib extension).  In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero.  The Reed librarians have created Zotero documentation at <http://libguides.reed.edu/citation/zotero>.  In addition, a tutorial is available from Middlebury College at <http://sites.middlebury.edu/zoteromiddlebury/>.

_R Markdown_ uses _pandoc_ (<http://pandoc.org/>) to build its bibliographies.  One nice caveat of this is that you won't have to do a second compile to load in references as standard \LaTeX\ requires. To cite references in your thesis (after creating your bibliography database), place the reference name inside square brackets and precede it by the "at" symbol.  For example, here's a reference to a book about worrying:  [@Molina1994].  This `Molina1994` entry appears in a file called `thesis.bib` in the `bib` folder.  This bibliography database file was created by a program called BibTeX.  You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the `bib` folder.  

For more information about BibTeX and bibliographies, see our CUS site (<http://web.reed.edu/cis/help/latex/index.html>)^[@reedweb2007]. There are three pages on this topic:  _bibtex_ (which talks about using BibTeX, at <http://web.reed.edu/cis/help/latex/bibtex.html>), _bibtexstyles_ (about how to find and use the bibliography style that best suits your needs, at <http://web.reed.edu/cis/help/latex/bibtexstyles.html>) and _bibman_ (which covers how to make and maintain a bibliography by hand, without BibTeX, at <http://web.reed.edu/cis/help/latex/bibman.html>). The last page will not be useful unless you have only a few sources.

If you look at the YAML header at the top of the main .Rmd file you can see that we can specify the style of the bibliography by referencing the appropriate csl file.  You can download a variety of different style files at <https://www.zotero.org/styles>.  Make sure to download the file into the csl folder.

##### Tips for Bibliographies

- Like with thesis formatting, the sooner you start compiling your bibliography for something as large as thesis, the better. Typing in source after source is mind-numbing enough; do you really want to do it for hours on end in late April? Think of it as procrastination.
- The cite key (a citation's label) needs to be unique from the other entries.
- When you have more than one author or editor, you need to separate each author's name by the word "and" e.g. `Author = {Noble, Sam and Youngberg, Jessica},`.
- Bibliographies made using BibTeX (whether manually or using a manager) accept \LaTeX\ markup, so you can italicize and add symbols as necessary.
- To force capitalization in an article title or where all lowercase is generally used, bracket the capital letter in curly braces.
- You can add a Reed Thesis citation^[@noble2002] option. The best way to do this is to use the phdthesis type of citation, and use the optional "type" field to enter "Reed thesis" or "Undergraduate thesis." 

## Anything else?

If you'd like to see examples of other things in this template, please contact the Data @ Reed team (email <data@reed.edu>) with your suggestions. We love to see people using _R Markdown_ for their theses, and are happy to help.

