---
output: pdf_document
---

# Conclusion {.unnumbered}
  \setcounter{chapter}{4}
	\setcounter{section}{0}

Idea de Rosario
-ver dell cluster de saxitoxin cuantos pasos se necesitron para llegar ahi.  
-A donde se iria el resultado de abrir el GMP  
-Otra vez, que Actinos tienen FolE  

If we don't want Conclusion to have a chapter number next to it, we can add the `{.unnumbered}` attribute.  This has an unintended consequence of the sections being labeled as 3.6 for example though instead of 4.1.  The \LaTeX\ commands immediately following the Conclusion declaration get things back on track.

#### More info

And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph. Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.

<!--
If you feel it necessary to include an appendix, it goes here.
-->

\appendix

# The First Appendix

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.

#### In the main Rmd file:

```{r ref.label = 'include_packages', results = 'hide', echo = TRUE}
```

#### In [](#ref_labels):

```{r ref.label = 'include_packages_2', results = 'hide', echo = TRUE}
```

# The Second Appendix, Open source code on this document
## R markdown  
Thanks to Rmakdown  Thesis  
Apendix one Useful docker commands   
-Create a new repository  
`docker build . -t evomining`  
`docker push nselemevomining`

## Docker  
Reinicie docker para liberar puertos  
`sudo service docker restart`   
`docker stop $(docker ps -a -q)`  

Detener todos los contenedores
`docker rm $(docker ps -a -q)`  

Remover contenedores detenidos
`docker rm $(docker ps -q -f status=exited)`

Remove all images
`docker rmi $(docker images -q)`  

Gblocks only runs inside folder /var/www/html/EvoMining  

lista contenedores  
`docker ps -a`  

uninstall docker from ubuntu (Fresh start)  
`sudo apt-get purge docker-engine`  
`sudo apt-get autoremove --purge docker-engine`  
`rm -rf /var/lib/docker`   # This deletes all images, containers, and volumes

` docker run -i -t -v /home/nelly/GIT/EvoMining/:/var/www/html -p 80:80 newevomining /bin/bash`
`perl startevomining`  

## Git  
`git add --all`
`git commit -m "Some message"`  
`git push -u origin master`  
`git clone `

## Connect GitHub and DockerHub
automated builds
The Dockerfile is available to anyone with access to your Docker Hub repository.
Your repository is kept up-to-date with code changes automatically.


## Additional resources

- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>

- _R Markdown_ Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>

- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html>

- `ggplot2` Documentation - <http://docs.ggplot2.org/current/>
