---
output: pdf_document
---

# Conclusion {.unnumbered}
  \setcounter{chapter}{4}
	\setcounter{section}{0}

Idea de Rosario
-ver dell cluster de saxitoxin cuantos pasos se necesitron para llegar ahi.  
-A donde se iria el resultado de abrir el GMP  
-Otra vez, que Actinos tienen FolE  

If we don't want Conclusion to have a chapter number next to it, we can add the `{.unnumbered}` attribute.  This has an unintended consequence of the sections being labeled as 3.6 for example though instead of 4.1.  The \LaTeX\ commands immediately following the Conclusion declaration get things back on track.

#### More info

And here's some other random info: the first paragraph after a chapter title or section head _shouldn't be_ indented, because indents are to tell the reader that you're starting a new paragraph. Since that's obvious after a chapter or section title, proper typesetting doesn't add an indent there.

<!--
If you feel it necessary to include an appendix, it goes here.
-->

\appendix

# The First Appendix

This first appendix includes all of the R chunks of code that were hidden throughout the document (using the `include = FALSE` chunk tag) to help with readibility and/or setup.

#### In the main Rmd file:

```{r ref.label = 'include_packages', results = 'hide', echo = TRUE}
```

#### In [](#ref_labels):

```{r ref.label = 'include_packages_2', results = 'hide', echo = TRUE}
```

# The Second Appendix, Open source code on this document
## R markdown  
Thanks to Rmakdown  Thesis  
Apendix one Useful docker commands   
-Create a new repository  
`docker build . -t evomining`  
`docker push nselemevomining`

## Docker  
Restart docker and free all ports  
`sudo service docker restart`   

liast containers  
`docker ps -a`  

ssh or bash into a running docker container  
`sudo docker exec -i -t romantic_brahmagupta /bin/bash`
`docker exec -it <mycontainer> bash`  


Stop all containers  
`docker rm $(docker ps -a -q)`  

Remove stopped containers  
`docker rm $(docker ps -q -f status=exited)`

Remove all images
`docker rmi $(docker images -q)`  

uninstall docker from ubuntu (Fresh start)  
`sudo apt-get purge docker-engine`  
`sudo apt-get autoremove --purge docker-engine`  
`rm -rf /var/lib/docker`   # This deletes all images, containers, and volumes

Run Evomining container using nselem/newevomining image  
` docker run -i -t -v /home/nelly/GIT/EvoMining/:/var/www/html -p 80:80 nselem/newevomining /bin/bash`
  
Start evomining inside this container   
`perl startevomining`  

Vizualice a tree  
`http://10.10.100.234/EvoMining/cgi-bin/color_tree.pl?9&&/var/www/html/EvoMining/exchange/CyanosBBH_MiBIG_DB.faa_CYANOS`
file 9.new must be on folder volume CyanosBBH_MiBIG_DB.faa_CYANOS  

Find a perl module  
`perl -MList::Util -e'print $_ . " => " . $INC{$_} . "\n" for keys %INC'`
EvoMining notes  
Gblocks only runs inside folder /var/www/html/EvoMining  

## Git  
`git add --all`  
`git commit -m "Some message"`   
`git push -u origin master`   
`git clone `  

## Connect GitHub and DockerHub
automated builds
The Dockerfile is available to anyone with access to your Docker Hub repository.
Your repository is kept up-to-date with code changes automatically.


## Additional resources

- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>

- _R Markdown_ Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>

- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html>

- `ggplot2` Documentation - <http://docs.ggplot2.org/current/>
