---
title: "Buffers"
author: "NellySelem"
date: "May 18, 2017"
output: pdf_document
---

##Buffers
  
### _Lysis Buffer_  
Volumes to prepare 50ml. Potassium phosphate $pH 7.5$ $10mM$ can also be used instead of Tris-HCL    
$$
\label{Reaction_Buffer_TrpF}
\begin{tabular}{ l l l l l}
\hline \\ [-1.5ex]
 Component              &    Stock   & Final               & volume &    Condition\\ [1ex]
\hline \\ [-1.5ex]
 Tris-HCL               &      1M    &     $10mM$          &    $500 \mu l$ & $pH 7.5$ \\ [1ex]
 EDTA                   &     500mM  &    $0.5mM$          &    $50  \mu l$ & pH 7.7\\ [1ex]
 NaCl                   &      1M    &    $50mM$           &    2.5 ml & -       \\ [1ex]
 Glycerol               & $50\%$ &     $5\%$    &             5 ml  &-       \\ [1ex]
 DTT                    &   324mM    &   $0.1mM$    &       `r 50*.1/324*1000` $\mu l$     &Fresh addition\\ [1ex]
 Mix protease inhibitor &        -   &   $0.5 \mu l by ml$&   $25 \mu l$  &   Fresh addition\\ [1ex]
Lysozyme                &$10\frac{mg}{ml}$ &        1000X  &  $50 \mu l$  &Fresh addition\\ [1ex]
$H_2O$&                               &                    &  41.95 ml          & \\ [1ex]
\hline
\end{tabular}
$$


*** 
  TRIS-HCL stock: Calculate TRIS-HCL grams at $1M$ on a final volume of $200 ml$  
$C_F=1M$ $V_F= 200ml$ = $V_F=0.2 L$ , $mw=121.14\frac{g}{mol}$  
  
$Molaridad\times Volumen$ = $\frac{gramos}{peso molecular}$  
 $1M\times 0.2L =\frac{x gramos}{ 121.14 \frac{g}{mol}}$  
x gramos = (1M)(0.2L) 121.14 $\frac{g}{mol}$=`r 1*0.2*121.14` g = `r 1000*0.01*0.2*121.14` mg   

**** 
  Calculate EDTA grams at $500mM$ on a final volum of $20 ml$  
$C_F=500mM$ $V_F= 200ml$,$C_F=.5M$  $V_F=0.02 L$,$mw=372.24\frac{g}{mol}$  
  
(Molaridad)(Volumen)= $\frac{gramos}{peso molecular}$  
(0.5M)(0.02L)=$\frac{x gramos}{ 372.24 \frac{g}{mol}}$  
x gramos = (0.5M)(0.02L) 372.24 $\frac{g}{mol}$ =`r .5*0.02*372.24` g= `r .5*0.02*372.24 *1000`mg   

 
  _To prepare 50ml from stocks:_  
Tris-HCL:  x1M=1000mM=(50ml)(10mM)-> x=`r 50*10/1000` ml=`r 50*10/1000*1000`$\mu l$    
EDTA:       x500mM=(50ml)-> 0.5mM x=`r 50*.5/500` ml  =`r 50*.5/500*1000` $\mu l$    
NaCl: x1M=1000mM=(50ml)50mM ->  x=`r 50*50/1000` ml  
Glycerol x50%= (50ml)5% ->x=`r 50*5/50`ml  
$H_2O$ -> `r  50- 50*5/50-50*50/1000-50*.5/500-50*10/1000`ml  
  _Fresh addition_
DTT x 324mM =(50ml)0.1mM ->x=`r 50*.1/324`= `r 50*.1/324*1000`ml  $\mu l$  
Mix protease inhibitor $.5\mu l$ by ml -> `r .5*50` $\mu l$  
lysosyme 10mg/ml 1000X on 50ml -> $50 \mu l$  

***
```{r VolBuferA,echo=FALSE}
VolBufferA=500 ##ml
TrisStock=1000 ##mM 1M
NaClStock=2000 ##mM 2M
NaClFinal=500 #mM 
TrisFinal=20 ##mM

VolBufferB=200 ##ml
ImidazolStock=1000 ##mM
ImidazolFinal=500 ##mM
ImidazolMW=68.08 ##g/mol
ImidazolVolStock=500 ##ml

```

### _Buffer A_  (nickel affinity purification)  
> Tris-HCL `r TrisFinal` mM or Potassium phosphate `r TrisFinal`mM     
> NaCl `r NaClFinal`mM (Buffer A available at $500mM$)      
Adjust $pH$ at$7.5$      
  

Volumes for `r VolBufferA` ml  

\label{Reaction_Buffer_A}
\begin{tabular}{ l l l l l}
\hline \\ [-1.5ex]
Reactive  & Stock Concentration & Final Concentration   & Volume & Condition\\ [1ex]
\hline \\ [-1.5ex]
\hline
Tris-HCL &  `r TrisStock/1000`M     & `r TrisFinal` mM  & `r TrisFinal*VolBufferA/TrisStock` ml &$pH 7.5$ \\  [1ex]
$NaCl$ &    `r NaClStock/1000`M   & `r NaClFinal`mM    & `r NaClFinal*VolBufferA/NaClStock` ml & \\  [1ex]
\hline
\end{tabular}

***

***

### _Buffer B_ (Nickel affinity purification)  
> Tris-HCL `r TrisFinal` mM or Potassium phosphate `r TrisFinal`mM     
> NaCl `r NaClFinal`mM (Buffer A available at $500mM$)      
> Imidazol `r ImidazolFinal`mM      

Adjust $pH$ at$7.5$      

Volumes for `r VolBufferB` ml  

\label{Reaction_Buffer_B}
\begin{tabular}{ l l l l l}
\hline \\ [-1.5ex]
Reactive  & Stock Concentration & Final Concentration   & Volume & Condition\\ [1ex]
\hline \\ [-1.5ex]
\hline
Tris-HCL &  `r TrisStock/1000`M     & `r TrisFinal` mM  & `r TrisFinal*VolBufferB/TrisStock` ml &$pH 7.5$ \\  [1ex]
NaCl &    `r NaClStock/1000`M   & `r NaClFinal`mM    & `r NaClFinal*VolBufferB/NaClStock` ml & \\  [1ex]
Imidazol &    `r ImidazolStock/1000`M   & `r ImidazolFinal`mM    & `r ImidazolFinal*VolBufferB/ImidazolStock` ml & \\  [1ex]
\hline
\end{tabular}
  
*Note* Potassium phosphate $pH 7.5$ $10mM$ or Hepes $50mM$ could have been used on above buffers without noticing protein quality or quantity variations.  

 ***

Imidazol `r ImidazolStock` mM      
> _To prepare `ImidazolVolStock` ml from stocks:_  
$C_F=$ `r ImidazolStock`mM $V_F=$ 50ml,  $mw=$ `r ImidazolMW` $\frac{g}{mol}$  
  
(Molaridad)(Volumen)= $\frac{gramos}{peso molecular}$  
(`r ImidazolStock/1000`M)(`r ImidazolVolStock/1000`L)=$\frac{x gramos}{ `r ImidazolMW` \frac{g}{mol}}$  

x gramos = (`r ImidazolStock/1000`M)(`r ImidazolVolStock/1000`L) `r ImidazolMW` $\frac{g}{mol}$ =

> `r (ImidazolStock/1000)*(ImidazolVolStock/1000)*ImidazolMW` g

\clearpage 

### _Reaction buffer TrpF assay_  
Pre reaction Buffer   
volumes to prepare 50ml of TrpF reaction buffer stock   

\label{Reaction_Buffer_TrpF}
\begin{tabular}{ l l l l l}
\hline \\ [-1.5ex]
Reactive  & Stock Concentration & Final Concentration   & Volume & Condition\\ [1ex]
\hline \\ [-1.5ex]
\hline
Tris-HCL &  1M     & $50 mM$  & `r 50*50/1000` ml &$pH 7.5$ \\  [1ex]
$MgCl_2$ &  1M     & $2mM$    & `r 50*2/1000` ml & \\  [1ex]
EDTA     &  500mM  & $0.5 mM$ & `r 50*0.5/500` ml &\\  [1ex]
\hline
\end{tabular}
$H_2O$ = `r 50-50*50/1000-50*2/1000-50*0.5/500`  ml  

  
Stocks:  
  
$MgCl_2$ stock at 1M on 10ml.    
$MgCl_2$ mw:95,211 g/mol  
$(.01L)(1M)=\frac{xg}{95.211\frac{g}{mol}}$  = `r .01*1*95.211` g = `r 1000*.01*1*95.211` mg

***  

Volumes to 50ml stock  
Tris-HCL: x 1M= x 1000mM =50ml $\times 50 mM$ -> x= `r 50*50/1000` ml  
$MgCl_2$: x 1M= x 1000mM= 50ml $\times 2mM$  -> x=  `r 50*2/1000` ml =  `r 1000*50*2/1000` $\mu l$  
EDTA:     x 500mM = 50ml $\times 0.5 mM$  -> x= `r 50*0.5/500` ml = `r 1000*50*0.5/500` $\mu l$     

***

Following reaction buffer components must be added at the moment of assay and stocks must be prepared using as solvent the _TrpF pre reaction buffer solution_ previously described, NOT water. This precaution is to avoid alterations on reactive concentrations.  This components must be stored at $-20^{\circ}C$  
  
At the moment of essay prepare reaction Buffer: 5 ml will be enough for 24 holes of $160\mu l$  

\label{Reaction_Buffer_TrpF}
\begin{tabular}{ l l l l}
\hline \\ [-1.5ex]
Reactive  & Stock Concentration & Final Concentration   & Volume on $5ml$\\ [1ex]
\hline \\ [-1.5ex]
\hline
Anthranilic acid & 33mM  & $0-200\mu M$   & see trpF assay\\  [1ex]
PRPP &   24mM    &  1mM   &    `r 1000*5*1/24` $\mu l$ \\  [1ex]
DTT     &  324mM & $0.4 mM$ & `r 1000*5*.4/324` $\mu l$\\  [1ex]

\hline
\end{tabular}
  
+this is the substrate that varies from 0 to $200\mu M$ store at $-20^{\circ}C$   
(Needs to be prepared but there is sigma dust available)  
  
***  
Stocks:  
AA stock $xmM$ $15ml$  $68.57mg$ (15ml are minimum to prepare solution otherwise is hard to obtain solution)
$(.015L)(xM)=\frac{68.57g}{137.14\frac{g}{mol}}$ 
```{r AA_100mM_5ml}
AA=68.57/(137.14*.015)
```
$x=(68.57)/(137.14*.015)g=0.06857g=68.57mg$  


-PRPP stock $xmM$ 10ml  
(Final concentration must be always 5 times molar when compared with anthranilic acid) store at $-20^{\circ}C$ Comes in small quantites, be careful.  

First stock $xmM$ $10ml$  
$(.01L)(xM)=\frac{120mg}{499.98\frac{g}{mol}}$ 
```{r PRPP_xmM_10ml_120mg}
PRPP=1000*.12/(499.98*.01)
```
> $x=\frac{(.12)(499.98)}{(.01)}M$ = `r PRPP`mM

***
  
-DTT stock $324mM$ (Final concentration 0.4 mM) store at $-20^{\circ}$  
1g on 20ml of reaction Buffer   
$(20ml)(xM)=\frac{1g}{154.253\frac{g}{mol}}$  
  $xM=\frac{1g}{154.253\frac{g}{mol}\times .02L}$    
 
```{r DTT_stock_1g_20ml}
DTT=1/(154.253*.02)
``` 
  $x=0.3241428M=324mM$  

***

Prepare 5ml of Reaction Buffer with DTT and PRPP  
DTT: $(5ml)(.4mM)=xml(324mM)$ ->  $DTT_f=$ `r 5*.4/324` ml = `r 1000*5*.4/324` $\mu l$.    
PRPP=$(5ml)(1mM)=xml(24mM)$ =  $PRPP_f=$ `r 5*1/24` ml = `r 1000*5*1/24` $\mu l$.    

\clearpage 
  
### _Reaction buffer proFAR assay_    
- TES $50mM$   (573mg p/50ml)  
- Ammonium bicarbonate $200mM$  (790.5mg p/50ml)  
- $H_2O$ miliQ 50ml  
- Additionally HisF must be added fresh to each pool on plate   
 
   ***
   
HisF $2.5mM$ by pool  ($10\mu l$ by pool on a stock at 40 $\mu M$)  
  
On Final volume of 160 $\mu l$ by pool $x\mu l=\frac{(2.5mM)(160\mu l)}{40mM}=10\mu l$  
 $100mM$
 
   ***
  
Calculate TES grams at $50mM$ on a final volum of $50 ml$  
$C_F=50mM$ $V_F= 50ml$  
$C_F=0.05M$  $V_F=0.05 L$  
$mw=229.25\frac{g}{mol}$  

(Molaridad)(Volumen)= $\frac{gramos}{peso molecular}$  
(0.05M)(0.05L)=$\frac{x gramos}{ 229.25 \frac{g}{mol}}$  
x gramos = (0.05M)(0.05L) 229.25 $\frac{g}{mol}$   
x=0.573g=573mg  

   ***
  
Calculate ammonium bicarbonate grams at $200mM$ on a final volume of $50 ml$   
$C_F=200mM$ $V_F= 50ml$  
$C_F=0.2M$  $V_F=0.05 L$  
$mw=79.056\frac{g}{mol}$  `r 50*.1/364`
  
(Molaridad)(Volumen)= $\frac{gramos}{peso molecular}$  
(0.2M)(0.05L)=$\frac{x gramos}{ 79.056 \frac{g}{mol}}$  
x gramos = (0.2M)(0.05L) 79.056 $\frac{g}{mol}$   
x=0.7905g=790.5mg  

\clearpage 

  
  