---
title: "ProFARassay"
author: "NellySelem"
date: "May 18, 2017"
output: pdf_document
---

```{r include_reedtemplates_proFAR, include = FALSE}
# This chunk ensures that the reedtemplates package is installed and loaded
# This reedtemplates package includes the template files for the thesis and also
# two functions used for labeling and referencing
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")

if(!require(reedtemplates)){
  library(devtools)
  devtools::install_github("ismayc/reedtemplates")
  }
library(reedtemplates)
```

```{r load_pkgs_proFAR, message = FALSE,echo=FALSE}
# List of packages required for this analysis
pkg <- c("dplyr", "ggplot2", "knitr", "devtools","RColorBrewer")
# Check if packages are not installed and assign the
# names of the packages not installed to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]
# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg))
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
# Load packages
library(dplyr)
library(plyr)
library(reshape) ## melt
library(ggplot2)
library(knitr)
library(RColorBrewer)
hm.palette <- colorRampPalette(rev(brewer.pal(11, 'Spectral')), space='Lab')  
library(genstats) ## Next libraries are for coursera
library(devtools)
library(Biobase)
library(scales) # 
library(xlsx) # For save data on excel
library(knitr)
library(drc) # for fitting Michaelis Menten model
library(ggplot2) # for drawing

```

## ProFAR Enzymatic assay 
-Essay perform on clear box  
This is a fluorimetric assay, that measures proFAR dissapering and AICAR forming. Meassureas are taken on AICAR at 300nm. ProFAR extintion coefficient: 6069 L/M *cm a 300 nm  
Aldo: proFAR is an accoplated essay with tHisF enzyme (_Thermothoga maritima_) cloned on pET-15b.   
Reaction suggested by Aldo:     
24 $\mu M$ ProFAR  
$0.36 \mu M$ SubHisA $pH 7.5$   
HisF excess $2.5 \mu M$   

Substrate concentrations  
Total volume = $160 \mu l$  
Enzyme concentration according to substrate  
HisF 40 uM each/hole 5 $\mu l$   
To obtain kinetic parameters Beer-Lambert law must be used on all data: Absorbance equal to extintion coefficient (L/M*cm) times pole large on cm times solute concentration (M/L)  
  
A= e*l*C  

Once data are transformed from absorbance to concentration units, initial velocity $V_0$ can be computed (Concentration/time). For each ProFAR concentration an initial velocity will be obtained and Michaleis Menden parameters can be used. Pole longitud l will be considered as 1 on a fix volume of 160ul (Ernesto) (160 ul) Lianet  

$$
\centering
\label{Sco_jueves}
\begin{tabular}{c c c c}
\hline \\ [-1.5ex]
Concentration  & Buffer $(\mu l)$ & Enzyme $(\mu l)$   & proFAR $(\mu l)$\\ [1ex]
\hline \\ [-1.5ex]
\hline
 0 & 155.4  & 4.6 & 0   \\  [1ex]
   & 152.2  & 4.6 & 3.2 \\  [1ex]
   &  149   & 4.6 & 6.4 \\  [1ex]
   &  142.6 & 4.6 & 12.8\\  [1ex]
   & 139.4  & 4.6 & 16   \\  [1ex]
   & 136.2  & 4.6 & 19.2 \\  [1ex]
   & 131.4  & 4.6 & 24 \\  [1ex]
   & 123.4  & 4.6 & 32 *** 
\\  [1ex]
\hline
\end{tabular}
$$

$$
\centering
\label{Sco_viernes}
\begin{tabular}{c c c c}
\hline \\ [-1.5ex]
Concentration  & Buffer $(\mu l)$ & Enzyme $(\mu l)$   & proFAR $(\mu l)$\\ [1ex]
\hline \\ [-1.5ex]
\hline
 0 & 156.1  & 3.9 & 0   \\  [1ex]
   & 152.9  & 3.9 & 3.2 \\  [1ex]
   & 149.7  & 3.9 & 6.4 \\  [1ex]
   & 143.3 & 3.9 & 12.8\\  [1ex]
   & 140.1  & 3.9 & 16   \\  [1ex]
   & 136.9  & 3.9 & 19.2 \\  [1ex]
   & 132.1  & 3.9 & 24 \\  [1ex]
   & 124.1  & 3.9 & 32 \\  [1ex]
\hline
\end{tabular}
$$

$$
\centering
\label{ScoD11_A}Niquel 
\begin{tabular}{c c c c}
\hline \\ [-1.5ex]
Concentration  & Buffer $(\mu l)$ & Enzyme $(\mu l)$   & proFAR $(\mu l)$\\ [1ex]
\hline \\ [-1.5ex]
\hline
 0 & 154.6  & 5.4 & 0   \\  [1ex]
   & 151.4  & 5.4 & 3.2 \\  [1ex]
   & 148.2   & 5.4 & 6.4 \\  [1ex]
   & 141.8 & 5.4 & 12.8\\  [1ex]
   & 138.6  & 5.4 & 16   \\  [1ex]
   & 135.4  & 5.4 & 19.2 \\  [1ex]
   & 130.6  & 5.4 & 24 \\  [1ex]
   & 122.6  & 5.4 & 32 \\  [1ex]
\hline
\end{tabular}
$$
Buffer essay proFAR ammonium Bicarbonate 790.5 mg left: 200mM
TES 573 mg left a 50mM vF 50 ml
ProFARstock
\clearpage   
## Calculate your own proFAR data
```{r ProFAR_Stocks}
#PriA_stock=48.5 ##uM D11A
PriA_stock=40 ##uM Scoe
HisF_stock=40 ##uM Scoe

type="Scoe"
hole_vol=170 ##ul
#PriA_final=2 ##uM final concentration on hole
PriA_final=.020 ##uM = 20 nM final concentration on hole
HisF_final=1.25 ##uM = 20 nM final concentration on hole

PriA_vol=hole_vol*PriA_final/PriA_stock
PriA_Dil1=PriA_vol*10 ## Dil 1:10

HisF_vol=hole_vol*HisF_final/HisF_stock
PlateColum=2
## 
proFAR_stock=300000 ##300,000nM-> 500uM-> .5mM
#proFAR_stock=300 ##500,000nM-> 500uM-> .5mM
ProFAR_mw=577.331 #http://www.genome.jp/dbget-bin/www_bget?cpd:C04896
```

```{r ProFARStock}
vol_stock_ProFAR=316 ##ul `r vol_stock_ProFAR`
#con_stock_ProFAR=300 ##uM `r con_stock_ProFAR`
con_stock_ProFAR=300 ##uM `r con_stock_ProFAR`
```

Now I will calculate ProFAR g needed to obtain a final concentration of $`r con_stock_ProFAR` \mu M$ on a final volume of $`r vol_stock_ProFAR`\mu l$ 
$Molaridad\times Volumen$ = $\frac{gramos}{peso molecular}$  

$(\frac{`r con_stock_ProFAR`}{10^6}M\times\frac{`r vol_stock_ProFAR`}{10^6}L)=\frac{x gramos}{ `r ProFAR_mw` \frac{g}{mol}}$   

x gramos = $\frac{`r con_stock_ProFAR`}{10^6}M\frac{`r vol_stock_ProFAR`}{10^6}L`r ProFAR_mw`\frac{g}{mol}$

So ProFAR quantity needed is

$`r ProFAR_mw*con_stock_ProFAR*vol_stock_ProFAR/10**(12)` g =`r (1000*ProFAR_mw*con_stock_ProFAR*vol_stock_ProFAR)/10**(12)` mg$ 

ProFAR must be diluted on ProFAR Reaction Buffer 1X

***
```{r proFAR_tables}

PriA<-rep(PriA_Dil1,8)  ## 
HisF<-rep(HisF_vol,8) 
# Concentration between 0 -100 uM
proFARdata <- read.table(header=TRUE, text='
  proFAR_uM  
  100          
  75   
  60
  50
  40
  20
  10
  0  
')

#Calculating volumes according to data
proFARdata["Vol"] <-round(1000*hole_vol*proFARdata["proFAR_uM"]/proFAR_stock,digits=2) # That creates the new column named "MY_NEW_COLUMN" 

#proFARdata["PriA_Dil1_nM"]<-1000*round(PriA_final,digits=2)
proFARdata["PriA_ul"]<-round(PriA,digits=2)
#proFARdata["HisF_uM"]<-round(HisF_final,digits=2)
proFARdata["HisF_ul"]<-round(HisF,digits=2)

# As an example, the new column receives
proFARdata$Buffer <- round(hole_vol-proFARdata$Vol-proFARdata$PriA_ul,digits=2)

kable(proFARdata)
```

With PriA Dilution at 4uM to get a final concentration of  
With 
\clearpage  
  
  

##ProFAR synthesis (From Thesis Aldo)  
## Optimum ph determination for hisA activity  
## Determination of cinetic parameters  
## Todo list
Ask Brieba for 1g L-arabinose         Status: Not done
Check Resistance from TrpD and TrpC    Status: Done

## Other Reactives available
-Anthranilic acid (room temperature reactive)  
(Box ProFAR synthesis -20C)  
-D-ribose 5 phosphate disodium salt hidrate   
-D ribose 5 phosphate disodium salt hidrate  
-Phosphoenol piruvic acid  monosodium salt hidrate  
-5phospho Dribose 1 diphosphate pentasodium salt  
-piruvate kinase  
-Adenosin 5 triphosphate disodium salt hidrate  
-pyrophosphatase inorganic from ecoli    
-Adenosin 5 triphosphate ATP  

