---
title: "ProteinQuantification"
author: "NellySelem"
date: "May 18, 2017"
output: pdf_document
---

```{r include_reedtemplates_PQ, include = FALSE}
# This chunk ensures that the reedtemplates package is installed and loaded
# This reedtemplates package includes the template files for the thesis and also
# two functions used for labeling and referencing
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")

if(!require(reedtemplates)){
  library(devtools)
  devtools::install_github("ismayc/reedtemplates")
  }
library(reedtemplates)
```

```{r load_pkgs_chapterPQ, message = FALSE,echo=FALSE}
# List of packages required for this analysis
pkg <- c("dplyr", "ggplot2", "knitr", "devtools","RColorBrewer")
# Check if packages are not installed and assign the
# names of the packages not installed to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]
# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg))
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
# Load packages
library(dplyr)
library(plyr)
library(reshape) ## melt
library(ggplot2)
library(knitr)
library(RColorBrewer)
hm.palette <- colorRampPalette(rev(brewer.pal(11, 'Spectral')), space='Lab')  
library(genstats) ## Next libraries are for coursera
library(devtools)
library(Biobase)
library(scales) # 
library(xlsx) # For save data on excel
library(knitr)
library(drc) # for fitting Michaelis Menten model
library(ggplot2) # for drawing

```

## Protein Quantification
### Bradford calibration curve  
Mix well while pipeting. Holes with protein content must become blue.  
Incubate 5 min.  
Measure at 595 abs (TECAN). clear box.  
5 ul y 50ul of protein. Take measures inside the curve range.  
BSA on ug/ml

$$ \label{ScoD11_A} 
\begin{tabular}{c c c c}
\hline 
BSA $\mu l$ por ml & BSA $\mu l$ & Bradford $\mu l$ & Other \\ [1ex]
\hline \\ [-1.5ex]  
0                & 0      & 40 & $H_2O$ 160 $\mu l$   \\ [1 ex]
10               & 160    & 40 &    -                 \\ [1 ex]
20               & 160    & 40 & -                    \\ [1 ex]
40               & 160    & 40 & -                    \\ [1 ex]
60               & 160    & 40 &-                     \\ [1 ex]
100              & 160    & 40 & -                    \\ [1 ex]
0                &   0    & 40 &  PriA 5 $\mu l$   $H_2O$ 155 $\mu l$   \\ [1 ex]
0                &   0    & 40 &  PriA 20 $\mu l$   $H_2O$ 130 $\mu l$  \\ [1 ex]

\hline  
\end {tabular}
$$




```{r BSA_Bradford}
## 160 ul of BSA on this concentrations on ug/ml were added to a 40 ul of Bradford
##BSA concentration on ug/ml
BSAConcentrationStock<-c(0,10,20,40,60,100)

#adding .16 ml on plates  
vol<-160
Bradford_df = data.frame(BSAConcentrationStock,vol)   

## quantity of microgrames on .16 ml
Bradford_df["Quantity"]<-BSAConcentrationStock*.160
## concentration after added 40 ul bradford (total 200 ul)
Bradford_df["Concentration200"]<-Bradford_df["Quantity"]/.2


## This are the absorbance measures for the different BSA concentrations  
Abs1BSA<-c(0.2649999857,0.4329000115,0.6754000187,0.9449999928,1.24240005,1.432000041)
Abs2BSA<-c(0.2948000133,0.4298999906,0.7348999977,0.9965000153,1.225999951,1.494899988)

Bradford_df["Abs1BSA"]<-Abs1BSA
Bradford_df["Abs2BSA"]<-Abs2BSA

## mean value of measures
Bradford_df$mean<-rowMeans(subset(Bradford_df, select = c(Abs1BSA, Abs2BSA)), na.rm = TRUE)

#adjusting blank
Bradford_df$meanAdjusted<-Bradford_df$mean-Bradford_df$mean[1]
BSA.lm = lm(meanAdjusted ~ Concentration200, data=Bradford_df) 
coeffs = coefficients(BSA.lm); coeffs 

b=coeffs[1]
m=coeffs[2]

eq = function(x){x*m+b}
plot(Bradford_df$Concentration200,Bradford_df$meanAdjusted,type="l")
par(new=TRUE)
plot(eq, 1,100)
```


### Enzyme quantification
For Bradford to work properly After protein concentration, 900 ul Buffer A (free of imidazol) must be added to 100 $\mu l$ sample

```{r PriA_measures}
##  ul of 
vol=0
type="Scoe"
ABS1=0.2606000006
ABS2=0.2978000045
ABS3=0.3059999943	
ABS4=0.260800004
ABS5=0.2989000082	
ABS6=0.3073999882
Pria_df=data.frame(vol,type,ABS1,ABS2,ABS3,ABS4,ABS5,ABS6)

Pria_df$type<-as.character(Pria_df$type)
data2=c(1,"Scoe",0.5029000044,0.4781999886,0.4738000035,0.5042999983,0.4801000059,0.4790999889)
data3=c(2,"Scoe",0.6236000061,0.6475999951,0.591899991,0.6348000169,0.6485000253,0.594900012)
data4=c(5,"Scoe",0.769299984,0.7829999924,0.769299984,0.7829999924,0.769299984,0.7829999924)
data5=c(10,"Scoe",1.2418999672,1.0677000284,1.2441999912,1.0859999657,1.2418999672,1.0677000284)
data6=c(10,"D11A",0.575600028,	0.576099992,0.575600028,	0.576099992,0.575600028,	0.576099992)
data7=c(20,"D11A",1.039499998,	1.067399979,1.039499998,	1.067399979,1.039499998,	1.067399979)

## add data to dataframe Pria_df
Pria_df=rbind(Pria_df,data2)
Pria_df=rbind(Pria_df,data3)
Pria_df=rbind(Pria_df,data4)
Pria_df=rbind(Pria_df,data5)
#Pria_df=rbind(Pria_df,data6)
#Pria_df=rbind(Pria_df,data7)

Pria_df$mean<-rowMeans(cbind(as.numeric(Pria_df$ABS1),as.numeric(Pria_df$ABS2),as.numeric(Pria_df$ABS3),as.numeric(Pria_df$ABS4),as.numeric(Pria_df$ABS5),as.numeric(Pria_df$ABS6)))

Pria_df$meanAdjusted=Pria_df$mean-Pria_df$mean[1]

## Pasando absorbancia a concentration de proteina en 200 ul (ug/ml)
## kari values 
#m=0.004679534923
#b=0.103674687
m
b

inter = function(y){(y-b)/m}
### interpolation leaves ug/ml
Pria_df$concentration200<-inter(Pria_df$meanAdjusted)

## protein quantity on ug on 200 ul (200ul =.2 ml)
Pria_df["Quantity"]=Pria_df$concentration200*.2
## that is the protein quantity that was additioned from the begining, in consecuence we calculate initial concentration 

### Inicial Concentration  =Quantity/Vol_inicial  ug/ul  ->1000 ug/ml
Pria_df["InitialConcentration"]=  1000*Pria_df$Quantity/as.numeric(Pria_df$vol)

showPriA=subset(Pria_df, select = -c(ABS1,ABS2,ABS3,ABS4,ABS5,ABS6,mean) )
kable(Bradford_df)
kable(showPriA)

```

Since 1kDa = 1Kg/mol and PriA Scoe molecular weight  is 25.7kDa we have that  
$M\times V = \frac{g}{mw}$  
$M=\frac{g}{V}\times\frac{1}{mw}$ 

```{r molarity}
## mas on Kda = kg/mol  
mw<-25.7 
# converting to g/mol  
mwgram<-mw*1000  
## concentration on g/L (after convert ug/ml)
concentScoe<-1.013
concentD11A<-1.248

concentScoe
concentD11A

## Concentration on uM (10**6M)
molar = function(x){(10**6)*x/mwgram}
molar(concentScoe)
molar(concentD11A)
```
## Concentration on microMolar
> So Scoe concentration is `r molar(concentScoe)`  and D11A is  `r molar(concentD11A)` $\mu M$

```{r adjustingM}

## we want final concentration 2.4uM on a final volume of 160 ul
## ciVi=CfVf
## Vi =CfVf/Vi
## Cf= 80uM Vf=5ul Vi=?ul Ci=x

adjustMol = function(x){2.5*160/x} 
## this are the ul need from stock cristal enzyme that will be diluted on buffer to complete 160 ul 
adjustMol(molar(concentScoe))
adjustMol(molar(concentD11A))
```

